---
title: 'lightGBM-R'
author: 'Peter Hahn'
date: '`r Sys.Date()`'
output:
  html_document:
    number_sections: true
    toc: true
---

# LightGBM for R
Basic script for implimentation of LightGBM including metrics for this competition
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(lightgbm)
library(GGally)
library(MLmetrics)
library(arrow)
library(rsample)
```

## Data loading and preprocessing
Feel free to experiment with the features. Only part of them are included in the final model
```{r loadtrainbook}
set.seed(123)
vol_df <- tibble(time_id = numeric(),
                 volat = numeric(),
                 volat2 = numeric(),
                 sd_bp1 = numeric(),
                 sd_ap1 = numeric(),
                 min_bp1 = numeric(),
                 min_ap1 = numeric(),
                 log_ret = numeric(),
                 log_ret2 = numeric(),
                 stock_id = numeric(),
                 price_spread = numeric(),
                 volume = numeric(),
                 imbalance = numeric()
                 )

## list with file names full.names and recursive to get all files within the directories
file_list <- list.files("./input/optiver-realized-volatility-prediction/book_train.parquet/", full.names = TRUE, all.files = FALSE, recursive = TRUE)
## list with names of directories. Used to convientently get the stock_id

file_names <- list.files("./input/optiver-realized-volatility-prediction/book_train.parquet/")

for (i in 1:length(file_list)){
        stock <- read_parquet(file_list[i])
        stock <- stock %>% mutate(wap= (bid_price1 * ask_size1 +ask_price1 * bid_size1) / (
                                       bid_size1 + ask_size1)) %>% 
                                        mutate(log_return = log(wap/lag(wap))) %>% 
                                        mutate(wap2= (bid_price2 * ask_size2 +ask_price2 * bid_size2) / (
                                       bid_size2 + ask_size2)) %>% 
                                        mutate(log_return2 = log(wap2/lag(wap2))) %>% 
                                        mutate(spread = ask_price1 - bid_price1) %>% 
                                        mutate(vol = ask_size1 + bid_size1) %>% 
                                        mutate(imb = ask_size1 - bid_size1)  %>% 
                                        group_by(time_id) %>%
                                        mutate(volat=sqrt(sum(diff(log((bid_price1 * ask_size1 + ask_price1 * bid_size1) 
                                                                        / (ask_size1 + bid_size1)))**2))) %>% 
                                        mutate(volat2=sqrt(sum(diff(log((bid_price2 * ask_size2 + ask_price2 * bid_size2) 
                                                                        / (ask_size2 + bid_size2)))**2))) %>%
                                                 summarize(sd_bp1 = mean(bid_price1),
                                                           min_bp1 = min(bid_price1),
                                                  sd_ap1 = sd(ask_price1),
                                                  min_ap1 = sd(ask_price1),
                                                  log_ret = mean(log_return),
                                                  log_ret2 = mean(log_return2),
                                                  volat = mean(volat),
                                                  volat2 = mean(volat2),
                                                  price_spread = mean(spread),
                                                  volume = mean(vol),
                                                  imbalance = mean(imb)
                                                  )

        ## add stock_id to rows
        stock <- stock %>% mutate(stock_id = as.numeric(str_sub(file_names[i],10)))
        vol_df <- vol_df %>% add_row(stock)
}
vol_df <- vol_df %>% filter(!is.na(log_ret)) %>% filter(!is.na(log_ret2))
```

### Load train and bind to features dataframe
```{r load-train}
train <- read_csv('../input/optiver-realized-volatility-prediction/train.csv')
vol_df <- left_join(vol_df, train ,by=c("stock_id","time_id"))
```

## Poly function 
```{r func_poly}
polyT=function(x,y){
polYi=matrix(0, length(x),3)
polYi[,1]=x
polYi[,2]=x**2
polYi[,3]=x**3    
colnames(polYi)=paste0(y, 1:3)
return(data.frame(polYi))
}
```

```{r}
poli <- polyT(vol_df$volat,"vola")
poli2 <- polyT(vol_df$volat2,"vola2" )
vol_df <- bind_cols(vol_df,poli,poli2)# %>% select(stock_id,target,var1,var2,var3)
```


If working locally save vol_df for further processing and reload by read
```{r eval = FALSE}
#write_csv(vol_df,"vol_df.csv")
vol_df <- read_csv("vol_df_main.csv")
```

## for first use only few variables
```{r}
vol_df_part <- vol_df %>% select(stock_id,time_id,target,log_ret,log_ret2,
                                 volat, volat2, volume, imbalance, price_spread, vola1, vola2, vola3 )
```


### strata on stock_id to get a good mix
```{r split-data}
set.seed(1234)
vol_df_split <- vol_df_part  %>% 
        initial_split(strata = stock_id)
        
df_train <- training(vol_df_split)
df_test <- testing(vol_df_split)
```
#LightGBM
## Prepare data for lgb
Building matrix and lgb dataset
```{r}
X_train <- df_train %>% select(-target) %>% as.matrix()
x_label <- df_train$target
Y_train <- df_test %>% select(-target) %>% as.matrix()
y_label <- df_test$target

vol_lgb <- lgb.Dataset(X_train, label = x_label, weight = 1/(x_label**2))
vol_lgb_test <- lgb.Dataset(Y_train, label = y_label, weight = 1/(y_label**2))
```

## define metrics functions
```{r}
lgb.RMSPE <- function(preds, vol_lgb){
    label <- getinfo(vol_lgb, "label")
    val <- sqrt(mean(((label - preds)/label)^2))
    return(list(name = "RMSPE", value = val, higher_better = FALSE))
    }
```

### grid parameters 
```{r}
lgb.grid = list(objective = "regression",
                boosting_type = "gbdt",
                metric = "lgb.RMSPE",
                min_sum_hessian_in_leaf = 1,
                feature_fraction = 0.7,
                bagging_fraction = 0.7,
                bagging_freq = 5,
                lambda_l2 = 1
        )
```

### cv- model 
```{r eval = FALSE}
lgb.model.cv <- lgb.cv(params = lgb.grid, data = vol_lgb, learning_rate = 0.02,  num_threads = 6, 
                       nrounds = 8000, early_stopping_rounds = 50, eval_freq = 20, eval = lgb.RMSPE, nfolds= n_folds)
```

### Final model with 
```{r}
lgb.model = lgb.train(params = lgb.grid,
                      data = vol_lgb,
                      learning_rate = 0.02,
                      num_leaves = 25,
                      num_threads = 2,
                      nrounds = 1081,
                      eval_freq = 20,
                      eval = lgb.RMSPE,
                      num_threads = 6)


```

# Prepare submission
```{r load-data}
set.seed(123)
vol_df_test <- tibble(time_id = numeric(),
                 volat = numeric(),
                 volat2 = numeric(),
                 sd_bp1 = numeric(),
                 sd_ap1 = numeric(),
                 min_bp1 = numeric(),
                 min_ap1 = numeric(),
                 log_ret = numeric(),
                 log_ret2 = numeric(),
                 stock_id = numeric(),
                 price_spread = numeric(),
                 volume = numeric(),
                 imbalance = numeric()
                 )

## list with file names full.names and recursive to get all files within the directories
file_list <- list.files("../input/optiver-realized-volatility-prediction/book_test.parquet/", 
                        full.names = TRUE, all.files = FALSE, recursive = TRUE)
## list with names of directories. Used to convientently get the stock_id

file_names <- list.files("../input/optiver-realized-volatility-prediction/book_test.parquet/")

for (i in 1:length(file_list)){
        stock <- read_parquet(file_list[i])
         stock <- stock %>% mutate(wap= (bid_price1 * ask_size1 +ask_price1 * bid_size1) / (
                                       bid_size1 + ask_size1)) %>% 
                                        mutate(log_return = log(wap/lag(wap))) %>% 
                                        mutate(wap2= (bid_price2 * ask_size2 +ask_price2 * bid_size2) / (
                                       bid_size2 + ask_size2)) %>% 
                                        mutate(log_return2 = log(wap2/lag(wap2))) %>% 
                                        mutate(spread = ask_price1 - bid_price1) %>% 
                                        mutate(vol = ask_size1 + bid_size1) %>% 
                                        mutate(imb = ask_size1 - bid_size1)  %>% 
                                        filter(!is.na(log_return)) %>% 
                                        filter(!is.na(log_return2))  %>%     
                                        group_by(time_id) %>%
                                        mutate(volat=sqrt(sum(diff(log((bid_price1 * ask_size1 + ask_price1 * bid_size1) 
                                                                        / (ask_size1 + bid_size1)))**2))) %>% 
                                        mutate(volat2=sqrt(sum(diff(log((bid_price2 * ask_size2 + ask_price2 * bid_size2) 
                                                                        / (ask_size2 + bid_size2)))**2))) %>%
                                                 summarize(sd_bp1 = mean(bid_price1),
                                                           min_bp1 = min(bid_price1),
                                                  sd_ap1 = sd(ask_price1),
                                                  min_ap1 = sd(ask_price1),
                                                  log_ret = mean(log_return),
                                                  log_ret2 = mean(log_return2),
                                                  volat = mean(volat),
                                                  volat2 = mean(volat2),
                                                  price_spread = mean(spread),
                                                  volume = mean(vol),
                                                  imbalance = mean(imb)
                                                  )
        ## add stock_id to rows
        stock <- stock %>% mutate(stock_id = as.numeric(str_sub(file_names[i],10)))
        vol_df_test <- vol_df_test %>% add_row(stock)
}
#vol_df_test <- vol_df_test %>% filter(!is.na(log_ret)) %>% filter(!is.na(log_ret2))
```

### standardize volume and imbalance

```{r}
vol_df_test <- vol_df_test %>% mutate(volume = (volume -mean(volume))/sd(volume)) %>% 
    mutate(imbalance = (imbalance - mean(imbalance))/sd(imbalance))
```

```{r}
poli <- polyT(vol_df_test$volat,"vola")
poli2 <- polyT(vol_df_test$volat2,"vola2" )
```

```{r}
vol_df_test <- bind_cols(vol_df_test,poli,poli2)# %>% select(stock_id,target,var1,var2,var3)
```

```{r}
vol_df_test <- vol_df_test %>% select(stock_id,time_id,log_ret,log_ret2,
                                 volat, volat2, volume, imbalance, price_spread, vola1, vola2, vola3 )
```

```{r make_prediction}
vol_df_mat <- vol_df_test %>% as.matrix()
pred <- predict(lgb.model, vol_df_mat)
vol_df_test$target <- pred 
vol_df_test <- vol_df_test %>% mutate(row_id = str_c(stock_id,"-",time_id))%>% select(row_id,target)
write_csv(vol_df_test,"submission.csv")
```

